---
title: "HW 1"
subtitle: "Data 180, Professor Kennedy"
author: 
  name: "Will Ferris"
  email: "ferrisw@dickinson.edu"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document
---
Make sure that you have accepted the hw invitation link, and working with your clone repository.

**Due date:** September 21 beginning of class.

**Reminder:** You can insert R code in your Rmd file as a chunk in the following way:
```{r}
# Your code goes here
```
Make sure to insert a chunk for each (sub)section of the assignment.

## Preliminary Task
Go to [datacamp.com](https://datacamp.com) and complete the tutorial "Introduction to R".


## Task 1
a) Store the data below in a data frame object named `my_df`.
```{r}
head(my_df)
person = c("Alice", "Bob", "Charlie", "Denise", "Erin", "Frank","Gisele","Harry")
preference = c("Twitter", "Facebook", "Twitter", "Instagram", "Facebook", "Twitter", "Twitter","Instagram")
my_df = data.frame(person, preference)
```


|Person | Preference|
|:------|:----------|
|Alice | Twitter |
| Bob | Facebook |
| Charlie | Twitter |
| Denise | Instagram |
| Erin | Facebook |
| Frank | Twitter |
| Gisele | Twitter |
| Harry | Instagram |

b) What are the dimensions of `my_df`?
- 8 rows and 2 columns
c) Write R code that counts the total number of users.
```{r}
nrow(my_df)
```

d) Write R code that counts the total number of users who prefer Twitter.
```{r}
my_df_twitter <- my_df[my_df$preference=="Twitter",]
my_df_twitter
nrow(my_df_twitter)
```

e) Write R code that returns what Gisele prefers.
```{r}
my_df_giselepref <- my_df[my_df$person=="Gisele",2]
my_df_giselepref
```

## Task 2
## a) Create a vector named `my_vec` that includes all integers from 1 to 10 (inclusive).
```{r}
my_vec<-c(1,2,3,4,5,6,7,8,9,10)
```

b) Replace the second element in `my_vec` with the number `99`.
```{r}
my_vec2<-c(1,99,3,4,5,6,7,8,9,10)
```

c) Write a function called `my_func` that prints `This element is greater than 5!` if an element in `myvec` is greater than 5, and `This element is NOT greater than 5!` otherwise.
```{r}
vec1_num<-c("1","2","3","4","5","6","7","8","9","10")
vec2_num<-c("1","99","3","4","5","6","7","8","9","10")
names(my_vec)<-vec1_num
names(my_vec2)<-vec2_num
my_func <- my_vec > 5:
  print("This element is greater than 5")
my_func2 <-my_vec2 >5:
  print("This element is greater than 5")
print(my_func)
print(my_func2)
else:
  print("this element is not greater than 5")
```

## Task 3
a. Write a function named `get_fibonacci` that returns the n'th value in the [fibonacci sequence](https://www.mathsisfun.com/numbers/fibonacci-sequence.html). For example, `get_fibonacci(5)` should return 3; `get_fibonacci(7)` should return 8 and so on. ***Hint:** It is possible to call the same function from inside it. E.g., you can call `get_fibonacci()` from inside `get_fibonacci()`. Like inception!*
```{r}
fib_vec<-c(0, 1, 1, 2, 3, 5, 8, 13, 21, 34, 55, 89, 144, 233, 377, 610, 987, 1597, 2584, 4181, 6765, 10946, 17711, 28657, 46368, 75025, 121393, 196418, 317811, 514229)
fib_num<-c("1","2","3","4","5","6","7","8","9","10","11","12","13","14","15","16","17","18","19","20","21","22","23","24","25","26","27","28","29","30")
names(fib_vec)<-fib_num
print(get_fibonacci)
```


b. Using your function in part (a), find and return the 19th value in the fibonacci sequence.
```{r}
get_fibonacci(19)
```


<br>
You are done! 🏁 Don't forget to `commit` and `push` your .Rmd file to your Github clone repository before the due date.
<br>

